---
title: 'Attractiveness vs performance'
author:
- name: Tomasz Przechlewski
  email: t.plata-przechlewski@psw.kwidzyn.edu.pl
  affiliation: Powiślańska Szkoła Wyższa (Kwidzyn/Poland)
date: "Dec. 2021"
output:
  html_document:
    includes:
      before_body: misc_hdr.html
description: (c) Tomasz Przechlewski / CC-BY license
---

```{r setup, include=FALSE}
library("dplyr")
library("tidyr")
knitr::opts_chunk$set(echo = TRUE)
```

# The problem

The data from the
*Attractiveness is positively related to World Cup performance in male, but not female, biathletes*
(https://academic.oup.com/beheco/article/30/5/1436/5525079)


Read the data (available from datadryad, cf 
 https://datadryad.org/search?utf8=%E2%9C%93&q=attractiveness+performance):

```{r}
b0 <- read.csv("biathlonDATA.csv", sep = ';',  header=T, na.string="NA" )
n0 <- nrow(b0)
```

Variables (some variables omitted):

* raterID: anonymous identifier for opposite-sex rater
* attractiveness: attractiveness rating on a scale on a scale from 1 (very unattractive) to 7 (very attractive)
* targetID: anonymous identifier for biathlete
* sex: biathlete's sex
* ethnicity: biathlete's ethnicity (coded as Cauc = Caucasian, nonCauc = non-Caucasian)
* zAge: biathlete's standardised age (z-score method)
* zHeight: biathlete's standardised height (z-score method)
* zBMI: biathlete's standardised body mass index (z-score method)
* zWCptsPB: biathletes' standardised career-best World Cup points score (z-score method)
* zAttractiveness: biathlete's standardised attractiveness rating (z-score method)
* meanAthlAttr: biathlete's mean standardised attractiveness rating (z-score method)

There are (`r n0`) observations

We transform the data: attractiveness is a sum of all ratings and the
other values are simply copied:

```{r}
b <- b0 %>% select (attractiveness, targetID, 
                    sex, ethnicity, zAge, zHeight, zBMI, zWCptsPB) %>%
  group_by(targetID) %>%
  summarise (att = sum(attractiveness, na.rm=T),
             id = first(targetID),
             sex = first(sex),
             eth=first(ethnicity),
             age = first(zAge),
             ht = first(zHeight),
             bmi = first(zBMI),
             p = first(zWCptsPB))
```

Check  the result:

```{r}
summary(b)
```

## Attractiveness vs Performance

```{r}
model1 <- lm(p ~ att, data = b)
summary(model1)

model2 <- lm(p ~ att + sex, data = b)
summary(model2)

model3 <- lm(p ~ age + eth, data = b)
summary(model3)

```

## Data banks

https://figshare.com/ (https://en.wikipedia.org/wiki/Figshare)

https://datadryad.org/ ( https://en.wikipedia.org/wiki/Dryad )

