---
title: "Milk production"
author: "TP"
date: "2022-01-01"
output:
   html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The aim is to forecast the monthly volume of milk production in Poland.
The source of the data is the Statistical Bulletin of the Central Statistical Office 10/2020;
https://stat.gov.pl/obszary-tematyczne/inne-opracowania/)

We use `forecast`/`fpp2` R packages (in version 8.13 or higher)


```{r echo=T}
library ("forecast")
library ("fpp2")

## show versions
sessionInfo()
```

The `milk.csv` file contains monthly data on milk production in Poland 
(January 2010--October 2020; table 47: purchase of major agricultural
products; https://stat.gov.pl/obszary-tematyczne/inne-opracowania/)


```{r echo=T}
## load data from csv / covert to TS 
t <- read.csv("mleko.csv", dec=",", sep = ';',  header=T, na.string="NA");
t <-ts(t, start=c(2010, 1), frequency=12)
## check last observation, should be 2020/10
end(t)
```

Visual recognition of the nature of the forecasted phenomenon (is the data complete,
whether there are any atypical observations. Is there a trend and seasonality.)

```{r echo=F}
## Try some of these
##  plot(t)
##  autoplot(t) 
##  ggtsdisplay(t, smooth = T)
autoplot(t, facets = TRUE) + geom_smooth()
```

Analyzing the chart, we can see clear seasonal fluctuations and a trend.
There are no missing data, unusual observations or trend breaks.
Another two charts
allow for a more accurate assessment of the seasonality and trend:

```{r echo=F}
ggseasonplot(t)
autoplot(decompose(t) )
```

The volume of milk purchases drops in winter and is the highest in the
months of May--July. The last chart clearly shows that  seasonality
of production
is additive (constant amplitude of fluctuations) and the
trend is close to linear

```{r echo=F}
## Ad 4/ statystyki zbiorcze
t.mean <- mean(t, na.rm = T)
t.max <- max(t, na.rm = T)
t.min <- min(t, na.rm = T)
#t.mean
#t.max
#t.min
```

In the period analyzed, the average milk production was `r t.mean` (minimum value` r t.min`; value
maximum `r t.max`)

## Prognostic assumptions

The **purchaser of the forecast** requested a forecast for 6 months.
Therefore, the length of the test set is set for a period of 6 months.


```{r echo=F}
## Test/train set 
## use window or (better)  head/tail
## tl <- head(t, round(length(t) * 0.8))
## h <- 99
## albo tl <- head(t, length(t) - h)
## tt <- tail(t, h)
hmax <- 6

tl <- head (t, length(t) -hmax)
tt <- tail (t, hmax)
cat ("End of train set:", end(tl))
cat ("End of test set:", start(tt))
```

## Linear trend forecasting

### Model estimation

```{r echo=F}
## 1. Modelu fit/summary
##    Assessing R2/RMSE/MAPE
m_lm <- tslm(tl ~ trend + season )
summary(m_lm)
accuracy(m_lm)
```

Evaluation of residuals

```{r echo=F}
checkresiduals(m_lm)
res_lm <- m_lm$residuals
```

```{r echo=F}
m_lm.fitted <- m_lm$fitted.values
autoplot(m_lm.fitted, series="teoret") +  autolayer(m_lm$x, series="empir")
# p > 0.05 świadczy o autokorelacji
Box.test(res_lm, type='Ljung-Box')
```

Conclusions: The model fits the data well (R> 90?) But it exists
autocorrelation of the random term.

Now we compute forecasts for next 6 months and compare with test set values:


```{r echo=F}
m_lmf <- forecast(m_lm, h=hmax)
autoplot(m_lmf)
m_mla <- accuracy(m_lmf, tt)
m_mla
```

## Forecasting with exponential smoothing

### Model estimation

```{r echo=F}
m_es  <- ets(tl)
summary(m_es)
accuracy(m_es)
```

Evaluation of residuals

```{r echo=F}
checkresiduals(m_es)
res_es <- m_es$residuals

m_es.fitted <- m_es$fitted
autoplot(m_es.fitted, series="teoret") +  autolayer(m_es$x, series="empir")

Box.test(res_es, type='Ljung-Box')
```

Now we compute forecasts for next 6 months and compare with test set values:

```{r echo=F}
m_esf <- forecast(m_es, h=hmax)
autoplot(m_esf)
m_esa <- accuracy(m_esf, tt)
m_esa
```

Conclusions:

## Arima

Model

```{r echo=F}
m_aa  <- auto.arima(tl)
summary(m_aa)
accuracy(m_aa)
```

Evaluation of residuals

```{r echo=F}
checkresiduals(m_aa)
res_aa <- m_aa$residuals

m_aa.fitted <- m_aa$fitted
autoplot(m_aa.fitted, series="teoret") +  autolayer(m_aa$x, series="empir")

Box.test(res_aa, type='Ljung-Box')
```

Now we compute forecasts for next 6 months and compare with test set values:

```{r echo=F}
m_aaf <- forecast(m_aa, h=hmax)
autoplot(m_aaf)
m_aaa <- accuracy(m_aaf, tt)
m_aaa
```

Conclusions

## Comparison of forecasts

As the series shows a trend and seasonality, we will compare the results
for *naive* snaive (seasonality) and rwf (trend) models

```{r echo=F}
m_sn  <- snaive(tl)
res_sn <- m_sn$residuals
m_sn.fitted <- m_sn$fitted
m_snf <- forecast(m_sn, h=hmax)
m_sna <- accuracy(m_snf, tt)
m_sna

m_rwf  <- rwf(tl)
res_rwd <- m_rwf$residuals
m_rwf.fitted <- m_rwf$fitted
m_rwff <- forecast(m_rwf, h=hmax)
m_rwfa <- accuracy(m_rwff, tt)
m_rwfa
```

Summary of fit scores (ordered by RMSE):

```{r echo=F}
## Najniższa wartość RMSE / MAPE w zbiorze testowym
## Poniższe przepisujemy po prostu
A.table <- rbind( m_mla,  m_esa, m_aaa, m_sna, m_rwfa)
row.names(A.table) <- c('lm', 'lm/t', 'es', 'es/t', 'arima', 'arima/t', 'sna', 'sna/t', 'rwf', 'ref/t')
A.table <- as.data.frame(A.table)
A.table <- A.table[order(A.table$RMSE),]
A.table
```

Using the lowest RMSE/MAPE value in the test file as the criterion
we choose the ARIMA model as the best predictor of milk purchase

Compare the forecasting results in the chart:


```{r echo=F}
autoplot(t, series="empir") +  
  autolayer(m_lmf$mean, series="ln") +
  autolayer(m_esf$mean, series="es") +
  autolayer(m_aaf$mean, series="aa") +
  autolayer(m_snf$mean, series="sn") +
  autolayer(m_rwf$mean, series="rwf")
 ```

 More detailed chart:
 
 ```{r echo=F}
## tylko zbiór testowy 
autoplot(tt, series="empir") +  
  autolayer(m_lmf$mean, series="ln") +
  autolayer(m_esf$mean, series="es") +
  autolayer(m_aaf$mean, series="aa")
```

All forecasts are systematically overestimated. Perhaps
this is due to the extraordinary decline
in purchasing related to the COVID19 pandemic.


